Practical Optional Types for Clojure

### 2016-01-07 intro.tex
add tejas, no table
explain ann, Any, Kw

The popularity of dynamically-typed languages in software
development, combined with a recognition that types often improve
programmer productivity, software reliability, and performance, has
led to the recent development of a wide variety of optional and
gradual type systems aimed at checking existing programs written in
existing languages.  These include  TypeScript~\cite{typescript} and Flow~\cite{flow} for
JavaScript, Hack~\cite{hack} for PHP, and mypy~\cite{mypy}
for Python among the optional systems, and Typed Racket~\cite{TF08}, Reticulated
Python~\cite{Vitousek14}, and GradualTalk~\cite{gradualtalk} among gradually-typed systems.\footnote{We
  use ``gradual typing'' for systems like Typed Racket with sound
  interoperation between typed and untyped code; Typed Clojure or
 TypeScript which don't
  enforce type invariants we describe as ``optionally typed''.}
%FIXME: add citations for all those systems.

@@

An example of these features is given in
\figref{fig:ex1}. Here, the \texttt{pname} multimethod dispatches
on the \texttt{class} of the argument---for \texttt{String}s,
the first method implementation is called, for \texttt{File}s, the
second. The \texttt{String} method calls
a \texttt{File} constructor, returning a non-nil \texttt{File} instance---the 
\texttt{.getName} method 
on \texttt{File} requires a non-nil target, returning a nilable
type.  
%Typed Clojure offers an opt-in mode that
%resolves JVM overloading, avoiding expensive runtime reflective calls.

>>>>

The popularity of dynamically-typed languages in software
development, combined with a recognition that types often improve
programmer productivity, software reliability, and performance, has
led to the recent development of a wide variety of optional and
gradual type systems aimed at checking existing programs written in
existing languages.  These include  TypeScript~\cite{typescript} and Flow~\cite{flow} for
JavaScript, Hack~\cite{hack} for PHP, and mypy~\cite{mypy}
for Python among the optional systems, and Typed Racket~\cite{TF08}, Reticulated
Python~\cite{Vitousek14}, and GradualTalk~\cite{gradualtalk} among gradually-typed systems.\footnote{We
  use ``gradual typing'' for systems like Typed Racket with sound
  interoperation between typed and untyped code; Typed Clojure or
 TypeScript which don't
  enforce type invariants we describe as ``optionally typed''.}

@@

An example of these features is given in
\figref{fig:ex1}. Here, the \clj{pname} multimethod dispatches
on the \clj{class} of the argument---for \clj{String}s,
the first method implementation is called, for \clj{File}s, the
second. The \clj{String} method calls
a \clj{File} constructor, returning a non-nil \clj{File} instance---the 
\clj{.getName} method 
on \clj{File} requires a non-nil target, returning a nilable
type.  
%Typed Clojure offers an opt-in mode that
%resolves JVM overloading, avoiding expensive runtime reflective calls.

### 2016-01-07 overview.tex
explain ann, Any, Kw

Here is a simple one-argument function
%
\clj{greet} that takes and returns strings.

@@

\begin{exmp}
\begin{minted}{clj}
(.getParent (new File "a/b"))  ;=> "a" :- (U nil Str)
\end{minted}
\label{example:getparent-direct-constructor}
\end{exmp}
%
Typed Clojure can integrate with the Clojure compiler to avoid expensive reflective 
calls like \clj{getParent}, however if a specific overload cannot be found based on the
surrounding static context, a type error is thrown.
%
\begin{minted}{clj}
(fn [f] (.getParent f)) ; Type Error: Unresolved interop: getParent
\end{minted}
%
Function arguments default to \clj{Any}, the most permissive type. Ascribing
a parameter type allows Typed Clojure to find a specific method.

@@

\paragraph{Value-based dispatch}
This simple multimethod says hello in different languages.%, as
%specified by a keyword argument.

>>>>

A simple one-argument function \clj{greet} is annotated with \clj{ann} to take and return strings.

@@

\begin{exmp}
\begin{minted}{clj}
(.getParent (new File "a/b"))  ;=> "a" :- (U nil Str)
\end{minted}
\label{example:getparent-direct-constructor}
\end{exmp}
%
Typed Clojure can integrate with the Clojure compiler to avoid expensive reflective 
calls like \clj{getParent}, however if a specific overload cannot be found based on the
surrounding static context, a type error is thrown.
%
\begin{minted}{clj}
(fn [f] (.getParent f)) ; Type Error: Unresolved interop: getParent
\end{minted}
%
Function arguments default to \clj{Any}, which is similar to a union of all types. Ascribing
a parameter type allows Typed Clojure to find a specific method.

@@

\paragraph{Value-based dispatch}
This simple multimethod takes a keyword (\clj{Kw}) and says hello in different languages.%, as
%specified by a keyword argument.
